---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

```{r}
library(tidyverse)
deaths <- av |>
  pivot_longer(
    c(Death1, Death2, Death3, Death4, Death5),
    names_to = "Time",
    values_to = "Death"
  ) |>
  mutate(
    Time = parse_number(Time),
    Death = as.factor(Death)
  )

deaths
```

Similarly, deal with the returns of characters.
```{r}
returns <- av |>
  pivot_longer(
    c(Return1, Return2, Return3, Return4, Return5),
    names_to = "Time",
    values_to = "Return"
  ) |>
  mutate(
    Time = parse_number(Time),
    Return = as.factor(Return)
  )

returns
```

Based on these datasets calculate the average number of deaths an Avenger suffers. 
```{r}
deaths_clean <- deaths |> 
  filter(Death != "")

deaths_clean |> 
  group_by(URL) |> 
  summarise(total_deaths = sum(Death == "YES")) |>
  summarise(average_deaths = mean(total_deaths))
```

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

Grace's Quote:

<<<<<<< HEAD
Jyotika's Quote: "I counted 89 total deaths — some unlucky Avengers (Like Mar-Vell and Hawkeye) are basically Meat Loaf with an E-ZPass — and on 57 occasions the individual made a comeback."
=======
“Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team. That’s about 40 percent of all people who have ever signed on to the team.”

Jyotika's Quote:
>>>>>>> a1a4743d42efd9d9ba8011e66e6b7ada980c4457

Alexander's Quote:

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

Grace's Code: 


```{r}
library(dplyr)

av %>%
  summarise(
    total_avengers = n(),
    total_died = sum(Death1 == "YES", na.rm = TRUE),
    percent_died = (total_died / total_avengers) * 100
  )

```





Jyotika's Code:
```{r}
total_deaths <- deaths |> 
  filter(Death == "YES") |> 
  summarise(count_deaths = n())

total_returns <- returns |> 
  filter(Return == "YES") |> 
  summarise(count_returns = n())

total_deaths
total_returns
```

Alexander's Code:

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.

Grace's Discussion: 

After running the code, I found that 69 of 173 Avengers have died at least once, giving a mortality rate of about 39.9%. This confirms FiveThirtyEight’s statement that approximately 40 percent of Avengers have died after joining the team. The numbers in the dataset align closely with the article’s findings, validating its analysis.

Jyotika's Discussion: 

It seems the total deaths are 145 and from this total deaths and there were 77 returns which means the article is incorrect in calculating the numbers. This quote is incorrect upon analysis.

Alexander's Discussion:


