---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

```{r}
library(tidyverse)
deaths <- av |>
  pivot_longer(
    c(Death1, Death2, Death3, Death4, Death5),
    names_to = "Time",
    values_to = "Death"
  ) |>
  mutate(
    Time = parse_number(Time),
    Death = as.factor(Death)
  )

deaths
```

Similarly, deal with the returns of characters.
```{r}
returns <- av |>
  pivot_longer(
    c(Return1, Return2, Return3, Return4, Return5),
    names_to = "Time",
    values_to = "Return"
  ) |>
  mutate(
    Time = parse_number(Time),
    Return = as.factor(Return)
  )

returns
```

Based on these datasets calculate the average number of deaths an Avenger suffers. 
```{r}
deaths_clean <- deaths |> 
  filter(Death != "")

deaths_clean |> 
  group_by(URL) |> 
  summarise(total_deaths = sum(Death == "YES")) |>
  summarise(average_deaths = mean(total_deaths))
```

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

Grace's Quote:

“Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team. That’s about 40 percent of all people who have ever signed on to the team.”

Jyotika's Quote:

"I counted 89 total deaths — some unlucky Avengers (Like Mar-Vell and Hawkeye) are basically Meat Loaf with an E-ZPass — and on 57 occasions the individual made a comeback."


Alexander's Quote:

"Given the Avengers’ 53 years in operation and overall mortality rate, fans of the comics can expect one current or former member to die every seven months or so"

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

Grace's Code: 


```{r}
library(dplyr)

av %>%
  summarise(
    total_avengers = n(),
    total_died = sum(Death1 == "YES", na.rm = TRUE),
    percent_died = (total_died / total_avengers) * 100
  )

```


Jyotika's Code:
```{r}
total_deaths <- deaths |> 
  filter(Death == "YES") |> 
  summarise(count_deaths = n())

total_returns <- returns |> 
  filter(Return == "YES") |> 
  summarise(count_returns = n())

total_deaths
total_returns
```

Alexander's Code:

```{r}
years = max(av$Years.since.joining)

# filter(av, Years.since.joining == 115)

# filter(av, Years.since.joining < 115)

dc <- deaths |>
  filter(Death != "")

dc2 <- dc |>
  group_by(URL) |>
  summarise(total_deaths = sum(Death == "YES"))

td <- sum(dc2$total_deaths)

mpd = (12 * years) / td

mpd53 = (12 * 53) / td

mpd53
```

Marcus' Code:

```{r}

library(tidyverse)
# install.packages("curl")

av1 <- read_csv(
  "https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv",
  show_col_types = FALSE
)

av2 <- av1 |>
  mutate(
    died_any = if_any(starts_with("Death"), ~ .x == "YES"),
    returned_any = if_any(starts_with("Return"), ~ .x == "YES")
  )

av2 |>
  group_by(`Current?`) |>
  summarise(
    p_died = mean(died_any, na.rm = TRUE),
    n = n()
  )

library(dplyr) 
library(scales)

summ <- av2 |>
  group_by(`Current?`) |>
  summarise(
    n = n(),
    died = sum(died_any, na.rm = TRUE),
    p_died = died / n,
    .groups = "drop"
  )

summ

prop.test(x = c(25, 44), n = c(82, 91))
```
### Tirtho’s Code

```{r Tirtho, echo=TRUE}
library(dplyr)

# Count total Avengers
total_avengers <- nrow(av)

# Count how many have died at least once
died_once <- av %>%
  mutate(died_any = if_any(starts_with("Death"), ~ .x == "YES")) %>%
  summarise(total_died = sum(died_any, na.rm = TRUE))

# Calculate percentage of Avengers who died at least once
percent_died <- (died_once$total_died / total_avengers) * 100

# Print results
total_avengers
died_once
percent_died

###My Thought:
After running the code, I found that 69 of 173 Avengers have died at least once, giving a mortality rate of about 40%.
This confirms FiveThirtyEight’s statement that approximately 40 percent of Avengers have died after joining the team.
The data supports the article’s claim and shows how many heroes in the dataset have been revived after their deaths.


### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.

Grace's Discussion: 

After running the code, I found that 69 of 173 Avengers have died at least once, giving a mortality rate of about 39.9%. This confirms FiveThirtyEight’s statement that approximately 40 percent of Avengers have died after joining the team. The numbers in the dataset align closely with the article’s findings, validating its analysis.

Jyotika's Discussion: 

I filtered and calculated the total deaths of the Avengers and then calculated the total returns of the individual. It seems the total deaths are 89 and there were 57 returns, which means upon analysis, the quote is correct.

Alexander's Discussion:

I first tested the assumption that the Avengers have been around for 53 years. I found the max value data set for Years Since Joining to be 115, meaning the Avengers has been around since at least 115 years ago. However, the next highest group is 53 years meaning there could be some problem with the data that explains these outliers. Finding the number of months before a death with 53 years resulted in 7, which is the number provided in the article. However, with the 115 years it is instead every 15 months. After testing it seems that they are likely correct with their assumptions if 53 years is total number of years the Avengers has been around instead of 115. 

Marcus' Discussion:

I tested whether current Avengers have fewer recorded deaths than former members. I marked a character as “died” if any of the death columns were “yes" Among current members, 25 of 82 (30.5%) have died at least once. Among non-current members, 44 of 91 (48.4%) have died. This is a 17.9 percentage difference, and a two-sample proportion test indicates it is statistically significant with p = 0.025. Therefore, the data support the claim that current Avengers have lower historical death rates than former members. The article summary reports that around 40% and the Avengers have died at least once and is often temporary so this checks out


